---
params:
  date_modified: ''
output:
  html_document:
    toc: yes
---

<style type="text/css">
.main-container {
  max-width: 1500px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{r setup, echo=F, warning=F, error=F, include=F}
knitr::opts_chunk$set(echo=F, warning=F, error=F)
library(here)
start.time <- Sys.time()
library(tidyverse)
library(gt)
library(glue)
library(gghsci)
library(ggplot2)
library(arrow)

p <- scales::label_number()
pp <- scales::label_percent()
if (is.null(getOption("arrow_duck_con"))) {
  options(
    arrow_duck_con = DBI::dbConnect(
      duckdb::duckdb(bigint = "integer64"),
      bigint = "integer64",
      dbdir = str_c(tempdir(), "/duckdb")
    )
  )
}
con <- getOption("arrow_duck_con")

DBI::dbExecute(con, "SET preserve_insertion_order = false;")
DBI::dbExecute(con, "SET memory_limit = '48GB';")

files <- Sys.glob(here("data/isni/*.parquet"))
dmtime <- if (params$date_modified!='') params$date_modified else strftime(fs::file_info(files[[1]])$modification_time,"%Y-%m-%d")

duckdb_open_parquet <- function(glob, db_connection=getOption("arrow_duck_con")) {
  tbl(db_connection, glue::glue("read_parquet('{glob}')"))
}

list.files(here("data/isni/")) |> 
        str_replace("(_[0-9])*\\.parquet","") |> 
        unique() |>
        walk(\(name) .GlobalEnv[[str_c('isni_',name)]] <- tbl(con, glue::glue("read_parquet('{here('data/isni')}/{name}*.parquet')")) |> pivot_longer(c(everything(), -isni_n), names_to="field") |>
  filter(!is.na(value)))

db_sample <- function(table, percent) {
  tbl(con, sql(glue::glue("{dbplyr::sql_render(table)} USING SAMPLE {percent}%")))
}

unit_print <- function(n) {
  return(case_when(
    n < 1000 ~ paste0(n, "B"),
    n < 1000000 ~ paste0(n / 1000, "KB"),
    n < 1000000000 ~ paste0(n / 1000000, "MB"),
    n < 1000000000000 ~ paste0(n / 1000000000, "GB"),
    n < 1000000000000000 ~ paste0(n / 1000000000000, "TB"),
    n < 1000000000000000000 ~ paste0(n / 1000000000000000, "PB"),
    TRUE ~ paste0(n / 1000000000000000000, "EB")))
}

n_isnis <- isni_core |> distinct(isni_n) |> count() |> pull() |> as.integer()
```

---
title: Composition Analysis of ISNI dated `r dmtime`
---

# Overall summary

`r p(n_isnis)` ISNIs in total. Constituent files:

```{r filelist, results = 'asis'}
cat("\n\n")
files[order(files |> str_replace(".parquet", ""))] |> 
  map(\(file) glue::glue("- {file} ({unit_print(fs::file_info(file)$size)})")) |>
  str_flatten(collapse="\n") |>
  cat()
cat("\n")
```

# Field usage overview

```{r field_usage_overview_data}
get_field_usage_stats <- function(dl) {
  field_example_vals <- dl |>
    count(field, value) |>
    group_by(field) |>
    slice_sample(n = 7) |>
    summarise(e_vals = str_flatten(str_c(value, ' (', format('{:t,}',n), ')'), collapse = "|"), .groups = "drop") |>
  #  union_all(subproperty_example_vals) |>
    collect()
  
  field_dist_vals <- dl |>
    group_by(field) |>
    summarise(dist_vals=n_distinct(value), .groups="drop") |>
    collect()
  
  field_usage_stats <- dl |>
    group_by(field) |>
    summarise(isnis = n_distinct(isni_n), n = n(), .groups = "drop") |>
    mutate(r_p = isnis / n_isnis) |>
    mutate(vals_per_rec = n / isnis) |>
    collect() |>
    left_join(
      dl |>
        count(isni_n, field) |>
        filter(n > 1L) |>
        ungroup() |>
        count(field, name = "isnis_with_mult_vals") |>
        collect(),
      join_by(field)
    ) |>
    replace_na(list(isnis_with_mult_vals = 0)) |>
    left_join(field_dist_vals, join_by(field)) |>
    left_join(field_example_vals, join_by(field)) |>
    relocate(field, isnis, r_p, dist_vals, e_vals) |>
    arrange(desc(isnis)) |>
    mutate(across(where(bit64::is.integer64), as.integer))
  rm(field_example_vals, field_dist_vals)
  field_usage_stats
}
```

Legend for the tables below:

* `isnis`: how many isnis have at least one of this field
* `r_p`: the above as a percentage of all isnis
* `dist_objs`: the number of different values appearing in this field
* `e_objs`: at most seven random sample values of this field
* `n`: the total number of appearances of this field
* `objs_per_rec`: calculated from the above, the mean number of appearances of this field per isni
* `subjs_with_mult_vals`: the number of isnis that have more than one repetition of this field

## Main table

```{r}
field_usage_stats <- get_field_usage_stats(isni_core)
```

In total, `r field_usage_stats |> distinct(field) |> count() |> pull(n) |> as.integer() |> p()` distinct fields are used.

```{r field_usage_overview_table}
field_usage_stats |>
  select(-c(vals_per_rec,isnis_with_mult_vals)) |>
  mutate(field = str_c('<a name="', field, '"></a>', field)) |>
  gt() |>
  fmt_number(where(is.numeric), decimals = 0) |>
  fmt_percent(r_p, decimals = 2, drop_trailing_zeros = TRUE) |>
  data_color(columns=r_p, palette="viridis") |>
  fmt_passthrough(field, escape=F)
```

## Other tables

```{r}
dl <- isni_names |>
  union_all(isni_same_as) |>
  union_all(isni_source_ids) |>
  union_all(isni_deprecated_isnis)
and_field_usage_stats <- get_field_usage_stats(dl)
```

In total, `r and_field_usage_stats |> distinct(field) |> count() |> pull(n) |> as.integer() |> p()` distinct fields are used.

```{r and_field_usage_overview_table}
and_field_usage_stats |>
  mutate(field = str_c('<a name="', field, '"></a>', field)) |>
#  select(-c(name, description)) |>
  gt() |>
  fmt_number(where(is.numeric), decimals = 0) |>
  fmt_percent(r_p, decimals = 2, drop_trailing_zeros = TRUE) |>
  data_color(columns=r_p, palette="viridis") |>
  data_color(columns=vals_per_rec, palette="viridis", method="quantile", quantiles=8) |>
  fmt_number(vals_per_rec, decimals = 2, drop_trailing_zeros = TRUE) |>
  fmt_passthrough(field, escape=F)
```

```{r time_taken}
end.time <- Sys.time()
time.taken <- end.time - start.time
#time.taken
```

Generation time: `r time.taken`
