---
params:
  date_modified: ''
output:
  html_document:
    toc: yes
---

<style type="text/css">
.main-container {
  max-width: 1500px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{r setup, echo=F, warning=F, error=F, include=F}
knitr::opts_chunk$set(echo=F, warning=F, error=F)
library(here)
start.time <- Sys.time()
library(tidyverse)
library(gt)
library(glue)
library(gghsci)
library(ggplot2)
library(arrow)

p <- scales::label_number()
pp <- scales::label_percent()
if (is.null(getOption("arrow_duck_con"))) {
  options(
    arrow_duck_con = DBI::dbConnect(
      duckdb::duckdb(bigint = "integer64"),
      bigint = "integer64",
      dbdir = str_c(tempdir(), "/duckdb")
    )
  )
}
con <- getOption("arrow_duck_con")

DBI::dbExecute(con, "SET preserve_insertion_order = false;")
DBI::dbExecute(con, "SET memory_limit = '48GB';")

files <- Sys.glob(here("data/wikidata/*.parquet"))
dmtime <- if (params$date_modified!='') params$date_modified else strftime(fs::file_info(files[[1]])$modification_time,"%Y-%m-%d")

duckdb_open_parquet <- function(glob, db_connection=getOption("arrow_duck_con")) {
  tbl(db_connection, glue::glue("read_parquet('{glob}')"))
}

list.files(here("data/wikidata/")) |> 
  keep(\(s) s|> str_detect(".parquet")) |>
  str_replace("(_[0-9])*\\.parquet","") |> 
  unique() |>
  walk(\(name) .GlobalEnv[[str_c('wd_',name |> str_replace_all("-","_"))]] <- tbl(con, glue::glue("read_parquet('{here('data/wikidata')}/{name}*.parquet')")))

db_sample <- function(table, percent) {
  tbl(con, sql(glue::glue("{dbplyr::sql_render(table)} USING SAMPLE {percent}%")))
}

kc <- function(content) {
  cat(sep = "\n", knitr::knit_child(quiet = TRUE, text = c(
    "```{r}",
    "content",
    "```"
  ), envir=environment()))
}

unit_print <- function(n) {
  return(case_when(
    n < 1000 ~ paste0(n, "B"),
    n < 1000000 ~ paste0(n / 1000, "KB"),
    n < 1000000000 ~ paste0(n / 1000000, "MB"),
    n < 1000000000000 ~ paste0(n / 1000000000, "GB"),
    n < 1000000000000000 ~ paste0(n / 1000000000000, "TB"),
    n < 1000000000000000000 ~ paste0(n / 1000000000000000, "PB"),
    TRUE ~ paste0(n / 1000000000000000000, "EB")))
}

p_instance_of <- wd_entities |> filter(id=="P31") |> pull(entity_id)
p_subclass_of <- wd_entities |> filter(id=="P279") |> pull(entity_id)

classes_of_interest <- list(human="Q5",corporate_body="Q106668099",written_work="Q47461344",edition="Q3331189") |> #work="Q386724"
  map(\(c_id) wd_entities |> filter(id==c_id) |> pull(entity_id))
  #organisation="Q43229",edition="Q3331189")#literary_work="Q7725634",written_work="Q47461344",film="Q11424",painting="Q3305213",human_settlement="Q486972")
#n_geonames <- d |> distinct(geonameid) |> count() |> pull() |> as.integer()
```

```{r}
wd_coreinfo <- wd_entities |>
  left_join(
    wd_labels |> 
      filter(language=="en") |>
      union_all(
        wd_labels |> 
          anti_join(wd_labels |> filter(language=="en"), join_by(entity_id)) |>
          group_by(entity_id) |>
          summarise(label=first(label))
      ),
    join_by(entity_id)
  ) |>
  transmute(entity_id, label=str_c(label, " (",id,")")) |>
  compute()
```

```{r}
wd_transitive_types <- function(class) {
  tbl(
    con,
    sql(glue::glue(
      "
  WITH RECURSIVE recursive_types AS (
    SELECT {class} AS type_entity_id
    UNION
    SELECT entity_id AS type_entity_id
    FROM ({dbplyr::sql_render(wd_claim_wikibase_entityid)})
    WHERE property_id == {p_subclass_of} AND rank!='deprecated' AND value_entity_id={class}
    UNION
    SELECT o.entity_id AS type_entity_id
    FROM recursive_types t, ({dbplyr::sql_render(wd_claim_wikibase_entityid)}) o
    WHERE o.property_id == {p_subclass_of} AND rank!='deprecated' AND t.type_entity_id == o.value_entity_id 
  )
  SELECT type_entity_id AS value_entity_id FROM recursive_types"
    ))
  )
}
```

```{r}
get_property_example_vals <- function(type_instances, claim_table, field) {
  field_name <- as_label(enquo(field))
  ret <- type_instances |>
    inner_join(claim_table, join_by(entity_id)) |>
    count(property_id, value={{field}}) |>
    group_by(property_id) |>
    slice_sample(n = 7)
  if (field_name |> str_ends("_id")) ret <- ret |>
    left_join(wd_coreinfo, join_by(value==entity_id)) |>
    mutate(value=label) |>
    select(-label)
  ret |>
    summarise(e_vals = str_flatten(str_c(value, ' (', format('{:t,}',n), ')'), collapse = "|"), .groups = "drop") |>
    collect()
}

get_property_dist_vals <- function(type_instances, claim_table, field) {
  type_instances |>
    inner_join(claim_table, join_by(entity_id)) |>
    group_by(property_id) |>
    summarise(dist_vals=n_distinct({{field}}), .groups="drop") |>
    collect()
}

get_property_usage_stats <- function(type_instances, n_instances, claim_table, field) {
  field_name <- as_label(enquo(field))
  type_instances |>
    inner_join(claim_table, join_by(entity_id)) |>
    group_by(property_id) |>
    summarise(subjects = n_distinct(entity_id), n = n(), .groups = "drop") |>
    mutate(r_p = subjects / n_instances) |>
    mutate(vals_per_subj = n / subjects) |>
    collect() |>
    left_join(
      type_instances |>
        inner_join(claim_table, join_by(entity_id)) |>
        count(entity_id, property_id) |>
        filter(n > 1L) |>
        ungroup() |>
        count(property_id, name = "subjs_with_mult_vals") |>
        collect(),
      join_by(property_id)
    ) |>
    replace_na(list(subjs_with_mult_vals = 0)) |>
    left_join(get_property_dist_vals(type_instances, claim_table, {{field}}), join_by(property_id)) |>
    left_join(get_property_example_vals(type_instances, claim_table, {{field}}), join_by(property_id)) |>
    left_join(
      type_instances |>
        inner_join(claim_table, join_by(entity_id)) |>
        distinct(property_id) |>
        left_join(wd_coreinfo |> 
          select(property_id=entity_id, property=label),
          join_by(property_id)
        ) |>
        collect()
      , join_by(property_id)) |>
    select(-property_id) |>
    mutate(field=field_name) |>
    relocate(property, field, subjects, r_p, dist_vals, e_vals) |>
    mutate(across(where(bit64::is.integer64), as.integer))
}
```

---
title: Composition Analysis of Wikidata dated `r dmtime`
---

# Overall summary

`r p(wd_coreinfo |> count() |> pull() |> as.integer())` entities in total. Constituent files:

```{r filelist, results = 'asis'}
cat("\n\n")
files[order(files |> str_replace(".parquet", ""))] |> 
  map(\(file) glue::glue("- {file} ({unit_print(fs::file_info(file)$size)})")) |>
  str_flatten(collapse="\n") |>
  cat()
cat("\n")
```

# Field usage overview

Legend for the tables below:

* `subjects`: how many subjects have at least one of this property
* `r_p`: the above as a percentage of all subjects
* `dist_vals`: the number of different values appearing in this property + field
* `e_vals`: at most seven random sample values of this property + field
* `n`: the total number of appearances of this property + field
* `vals_per_subj`: calculated from the above, the mean number of appearances of this property + field per subject
* `subjs_with_mult_vals`: the number of subjects that have more than one repetition of this property + field

```{r field_usage_overviews, results='asis'}
classes_of_interest |>
  walk(\(class) {
    glue::glue("\n\n## {wd_coreinfo |> filter(entity_id==class) |> pull(label)}\n\n") |> cat()

    type_instances <- wd_transitive_types(class) |>
      inner_join(wd_claim_wikibase_entityid |> filter(property_id==p_instance_of), join_by(value_entity_id)) |>
      distinct(entity_id) |>
      compute()

    n_type_instances <- type_instances |> count() |> pull()

    property_usage_stats <- list(
      get_property_usage_stats(type_instances, n_type_instances, wd_claim_globecoordinate, precision),
      get_property_usage_stats(type_instances, n_type_instances, wd_claim_globecoordinate, globe_entity_id),
      get_property_usage_stats(type_instances, n_type_instances, wd_claim_monolingualtext, language),
      get_property_usage_stats(type_instances, n_type_instances, wd_claim_monolingualtext |> mutate(language_text=str_c(language,": ", text)), language_text),
      get_property_usage_stats(type_instances, n_type_instances, wd_claim_string, value),
      get_property_usage_stats(type_instances, n_type_instances, wd_claim_wikibase_entityid, value_entity_id),
      get_property_usage_stats(type_instances, n_type_instances, wd_claim_time, timezone),
      get_property_usage_stats(type_instances, n_type_instances, wd_claim_time, before),
      get_property_usage_stats(type_instances, n_type_instances, wd_claim_time, after),
      get_property_usage_stats(type_instances, n_type_instances, wd_claim_time, precision),
      get_property_usage_stats(type_instances, n_type_instances, wd_claim_time, calendarmodel_entity_id),
      get_property_usage_stats(type_instances, n_type_instances, wd_claim_time, time)
    ) |>
      reduce(union_all) |>
      arrange(desc(subjects), property, field)
    glue::glue("In total, {property_usage_stats |> distinct(property) |> count() |> pull(n) |> as.integer() |> p()} distinct properties are used.") |> cat()
    property_usage_stats |>
      filter(subjects>n_type_instances/100/2) |>
#  mutate(property = str_c('<a name="', property, '"></a>', property)) |>
#  select(-c(name, description)) |>
      gt() |>
      fmt_number(where(is.numeric), decimals = 0) |>
      fmt_percent(r_p, decimals = 2, drop_trailing_zeros = TRUE) |>
      data_color(columns=r_p, palette="viridis") |>
    #  fmt_number(vals_per_rec, decimals = 2, drop_trailing_zeros = TRUE) |>
      #maybe_color() |>
      fmt_passthrough(property, escape=F) |>
      kc()
  })
#maybe_color <- \(x) if (field_usage_stats |> distinct(vals_per_rec) |> count() |> pull() == 1) x else x |> data_color(columns=vals_per_rec, palette="viridis", method="quantile", quantiles=8) # quantile has a bug when all same
```

```{r time_taken}
end.time <- Sys.time()
time.taken <- end.time - start.time
#time.taken
```

Generation time: `r time.taken`
