---
params: 
  dataset: ''
  dataset_name: ''
  property_descriptions: ''
  out: ''
  date_modified: ''
  help: ''
output:
  html_document:
    toc: yes
---

<style type="text/css">
.main-container {
  max-width: 1500px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{r setup, echo=F, warning=F, error=F, include=F}
knitr::opts_chunk$set(echo=F, warning=F, error=F)
library(here)
#params=list(dataset_name="ULAN",dataset=here("data/ulan/ulan*.parquet"),property_descriptions="../data/schema-info/rdf_properties.tsv", min_year=1400,max_year=1800, date_modified='')
start.time <- Sys.time()
library(tidyverse)
library(gt)
library(glue)
library(gghsci)
library(ggplot2)
library(arrow)

p <- scales::label_number()
pp <- scales::label_percent()
if (is.null(getOption("arrow_duck_con"))) {
  options(
    arrow_duck_con = DBI::dbConnect(
      duckdb::duckdb(bigint = "integer64"),
      bigint = "integer64",
      dbdir = str_c(tempdir(), "/duckdb")
    )
  )
}
con <- getOption("arrow_duck_con")

DBI::dbExecute(con, "SET preserve_insertion_order = false;")
DBI::dbExecute(con, "SET memory_limit = '48GB';")

files <- Sys.glob(params$dataset)
dmtime <- if (params$date_modified!='') params$date_modified else strftime(fs::file_info(files[[1]])$modification_time,"%Y-%m-%d")

duckdb_open_parquet <- function(glob, db_connection=getOption("arrow_duck_con")) {
  tbl(db_connection, glue::glue("read_parquet('{glob}')"))
}

d <- duckdb_open_parquet(params$dataset)

db_sample <- function(table, percent) {
  tbl(con, sql(glue::glue("{dbplyr::sql_render(table)} USING SAMPLE {percent}%")))
}

property_descriptions <- if (params$property_descriptions!='') read_tsv(params$property_descriptions) else tibble(
  property = character(),
  name = character(),
  description = character()
)

unit_print <- function(n) {
  return(case_when(
    n < 1000 ~ paste0(n, "B"),
    n < 1000000 ~ paste0(n / 1000, "KB"),
    n < 1000000000 ~ paste0(n / 1000000, "MB"),
    n < 1000000000000 ~ paste0(n / 1000000000, "GB"),
    n < 1000000000000000 ~ paste0(n / 1000000000000, "TB"),
    n < 1000000000000000000 ~ paste0(n / 1000000000000000, "PB"),
    TRUE ~ paste0(n / 1000000000000000000, "EB")))
}

types <- d |> 
  filter(property=="rdf:type") |> 
  group_by(subject) |> 
  summarise(types=sql("STRING_AGG(DISTINCT object, ', ' ORDER BY object)")) |> 
  compute()

subjects <- d |> 
  distinct(subject) |> 
  compute()
#n_subjects <- d |> distinct(subject) |> count() |> pull() |> as.integer()
```

---
title: Composition Analysis of the `r params$dataset_name` dated `r dmtime`
---

# Overall summary

`r p(subjects |> count() |> pull() |> as.integer())` subjects and `r p(d |> count() |> pull() |> as.integer())` triples in total. Constituent files:

```{r filelist, results = 'asis'}

cat("\n\n")
for (file in files[order(files |> str_replace(".parquet", ""))]) {
  unit_size <- unit_print(fs::file_info(file)$size)
  cat(glue::glue("- {file} ({unit_size})\n"))
}
cat("\n")
```

## Data types

```{r types}
types |>
  count(types) |>
  mutate(n=as.integer(n)) |>
  arrange(desc(n)) |>
  collect() |>
  gt(rowname_col="types") |>
  fmt_number(where(is.numeric), decimals = 0)
#  fmt_percent(r_p, decimals = 2, drop_trailing_zeros = TRUE) |>
#  fmt_number(vals_per_rec, decimals = 2, drop_trailing_zeros = TRUE) |>
#  data_color(columns=r_p, palette="viridis") |>
#  data_color(columns=vals_per_rec, palette="viridis", method="quantile", quantiles=8) |>
#  fmt_passthrough(subproperty, escape=F)
```

`r subjects |> anti_join(types, join_by(subject)) |> count() |> pull() |> as.integer() |> p()` subjects do not have type info.

# Property usage overview

```{r property_usage_overview_data}

property_example_objs <- d |>
  left_join(types, join_by(subject)) |>
  count(types, property, object) |>
  group_by(types, property) |>
  slice_sample(n = 7) |>
  summarise(e_objs = str_flatten(str_c(object, ' (', n, ')'), collapse = "|"), .groups = "drop") |>
#  union_all(subproperty_example_vals) |>
  collect()

property_dist_objs <- d |>
  left_join(types, join_by(subject)) |>
  group_by(types, property) |>
  summarise(dist_objs=n_distinct(object), .groups="drop") |>
  collect()

property_usage_stats <- d |>
  left_join(types |>
    inner_join(types |> count(types, name="n_subjects"), join_by(types)), join_by(subject)) |>
  replace_na(list(n_subjects=subjects |> count() |> pull() - types |> count() |> pull())) |>
  group_by(types, n_subjects, property) |>
  summarise(subjects = n_distinct(subject), n = n(), .groups = "drop") |>
  mutate(r_p = subjects / n_subjects) |>
  mutate(objs_per_rec = n / subjects) |>
  collect() |>
  left_join(
    d |>
      count(subject, property) |>
      filter(n > 1L) |>
      ungroup() |>
      left_join(types, join_by(subject)) |>
      count(types, property, name = "subjs_with_mult_vals") |>
      collect(),
    join_by(types, property)
  ) |>
  replace_na(list(subjs_with_mult_vals = 0)) |>
  left_join(property_dist_objs, join_by(types, property)) |>
  left_join(property_example_objs, join_by(types, property)) |>
  left_join(property_descriptions, join_by(property)) |>
  relocate(property, description, subjects, r_p, dist_objs, e_objs) |>
  arrange(property) |>
  mutate(across(where(bit64::is.integer64), as.integer))

rm(property_example_objs, property_dist_objs)
```

In total, `r property_usage_stats |> distinct(property) |> count() |> pull(n) |> as.integer() |> p()` distinct properties are used.

Legend for the tables below:

* `subjects`: how many subjects have at least one of this property
* `r_p`: the above as a percentage of all subjects
* `dist_objs`: the number of different values appearing in this property
* `e_objs`: at most seven random sample values of this property
* `n`: the total number of appearances of this property
* `objs_per_rec`: calculated from the above, the mean number of appearances of this property per subject
* `subjs_with_mult_vals`: the number of subjects that have more than one repetition of this property

```{r property_usage_overview_table}

property_usage_stats |>
  group_by(types) |>
  group_walk(\(t,s) {
    knitr::asis_output(glue::glue("## {s}")) |> print()
    maybe_color <- \(x) if (t |> distinct(objs_per_rec) |> count() |> pull() == 1) x else x |> data_color(columns=objs_per_rec, palette="viridis", method="quantile", quantiles=8) # quantile has a bug when all same
    t |>
  arrange(desc(subjects)) |>
  mutate(property = str_c('<a name="', property, '"></a>', property)) |>
#  select(-c(name, description)) |>
  gt() |>
  fmt_number(where(is.numeric), decimals = 0) |>
  fmt_percent(r_p, decimals = 2, drop_trailing_zeros = TRUE) |>
  data_color(columns=r_p, palette="viridis") |>
  fmt_number(objs_per_rec, decimals = 2, drop_trailing_zeros = TRUE) |>
  maybe_color() |>
  fmt_passthrough(property, escape=F) |> print()
  })
```

```{r time_taken}
end.time <- Sys.time()
time.taken <- end.time - start.time
#time.taken
```

